---
title: "M3: Class Exercise on Trend and Season - after class Jan 31"
author: "Haru Koga"
output: pdf_document
---

## Setting R code chunk options

First R code chunk is used for setting the options for all R code chunks. The choice echo=TRUE means both code and output will appear on report, include = FALSE neither code nor output is printed.

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output. 

```{r package, message=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)

```

## Importing data

Let's continue working with our inflow data for reservoirs in Brazil. 

```{r}

#Importing time series data from text file#
raw_inflow_data <- read.table(file="../Data/inflowtimeseries.txt",header=FALSE,skip=0)

#Trim the table to include only columns you need
nhydro <- ncol(raw_inflow_data)-2
nobs <- nrow(raw_inflow_data) 

#If your file does not have header like this one you can add column names after 
#creating the data frame
colnames(raw_inflow_data)=c("Month","Year", "HP1", "HP2","HP3","HP4", "HP5",
                            "HP6","HP7", "HP8","HP9","HP10", "HP11","HP12", 
                            "HP13", "HP14","HP15")

#Checking data
head(raw_inflow_data)
str(raw_inflow_data)
```
## Creating the date object

Here we use the function my() from package lubridate.

```{r}
#using package lubridate
my_date <- paste(raw_inflow_data[,1],raw_inflow_data[,2],sep="-")
my_date <- my(my_date)  #function my from package lubridate
head(my_date)
#add that to inflow_data and store in a new data frame 
inflow_data <- cbind(my_date,raw_inflow_data[,3:(3+nhydro-1)])
head(inflow_data)

```


## Initial Plots

Initial time series plot.

```{r}
#using package ggplot2
for(i in 1:nhydro){
  print(ggplot(inflow_data, aes(x=my_date, y=inflow_data[,(1+i)])) +
            geom_line(color="blue") +
            ylab(paste0("Inflow ",colnames(inflow_data)[(1+i)],sep="")) 
        )
}

```

## Zeros in the end on data 

The initial plots showed that we have zeros in the end of the data set. It could be missing observation or observation that haven't been observed yet. Use the tail() to find out how many zeros you have and how many lines you will need to remove.

```{r}
#check the final obs on data 
tail(inflow_data)

```

Note our last observation is from August 2011 but the data file was filled with zeros. Let's remove the last four rows of our data set.

```{r}
#Remove last year by replacing current data frame
inflow_data <- inflow_data[1:(nobs-12),]

#update object with number of observations
nobs <- nobs-12

#Tail again to check if the rows were correctly removed
tail(inflow_data)
```
Fixed!

## Transforming data into time series object

Many of the functions we will use require a time series object. You can transform your data in a time series using the function *ts()*. 

```{r}
ts_inflow_data <- ts(inflow_data[,2:(2+nhydro-1)],frequency=12,start=c(1931,1)) #year,month,date
#monthly data's frequency is 12

ts_inflow_data 

```


## Plotting ACF and PACF

Let's use functions Acf() and Pacf() from package "forecast".

```{r acf}

for(i in 1:nhydro){
  par(mfrow=c(1:2))
  Acf(ts_inflow_data[,i],lag.max=40,main=paste("HP",i),ylim=c(-1,1))
  Pacf(ts_inflow_data[,i],lag.max=40,ylim=c(-1,1))
}

#non stationarity 
#seasonality of the correlation 
#look at y-axix to see the magnetude of correlation 

#answer seasonality and magnitude of correlation just looking at y-axix
#HP4 declining - non-stationary
#HP8 weak autocorrelation 

```


## Trend Component

Let's identify and remove trend component like we leaned on the recorded videos for M4. You start by fitting a linear model to $Y_t=\beta_0+\beta_1*t+\epsilon_t$. 

```{r}
#Create vector t

t <- 1:nobs
iHP=4 #name of the index

#Fit a linear trend to TS of iHP
linear_trend <- lm(inflow_data[,iHP+1] ~ t)
summary(linear_trend)
# you can just say 5 to replace iHP+1

ggplot(inflow_data,aes(x=my_date, y=inflow_data[,iHP+1]))+
  geom_line(color="blue")+
  geom_smooth(method = "lm",color="red")

```
t 2.1836 = beta 1, the intercept is beta 0. 

```{r}
beta0 <- linear_trend$coefficients[1] #intercept 
beta1 <- linear_trend$coefficients[2] #beta 1 is not the mean but level of the series. 

#remove the trend from series
# you can use either TS object or dataframe.

detrend_inflow_data <- inflow_data[,(iHP+1)]- (beta0+beta1*t)

ggplot(inflow_data,aes(x=my_date,y=inflow_data[,iHP+1]))+
  geom_line(color="blue")+
  geom_line(aes(y=detrend_inflow_data),color="green")

```
Note that blue line is our original series, red lien is our trend, green line is our original series minus the trend or in other words the detrended series. And in orange is the trend line for the detrended series which has slope 0 meaning we were able to effectively eliminate the trend with a linear model.

## Seasonal Component

Now let's shift attention to the seasonal component.

```{r}
#Use seasonal means model
#by creating seasonal dummies and running a regression over the dummies 

iHP=1

dummies <- seasonaldummy(ts_inflow_data[,iHP]) # leave December out for intercept on regression 

seas_means_model <- lm(inflow_data[,(iHP+1)]~dummies)
summary(seas_means_model)
#negative month - bs it's relative to December. Intercept value is the mean for December. your x value is just showing seasonal components. 
#if you want the Jan mean, intercept + the coefficient on dummiesJan. 

beta0 <- seas_means_model$coefficients[1]
beta1 <- seas_means_model$coefficients[2:12]

inflow_seas_comp <- array(0,nobs)
for(i in 1:nobs){
  inflow_seas_comp[i] = beta0 + beta1 %*% dummies[i,]
} 
#effect of seasonal components are added

ggplot(inflow_data,aes(x=my_date,y=inflow_data[,iHP+1]))+ 
  geom_line(color="blue")+
  geom_line(aes(y=inflow_seas_comp),color="orange")
  
  
#use differencing at lag equal to period of dataset. to model a seasonal component. 
#when you have a short period of years, using this model is not a concern. bc the January mean of 1940 and 2000 are the same in this model.
#variation on seasonality and trend are different things. 

#go with TS object and model seasonal component
```

##Exercise

Fit trend and seasonal for the other variables HP2, HP3, ...


## Stationarity Tests in R

Some test only work for non-seasonal data. So let's create another series with yearly averages for inflow.

```{r}
#Group data in yearly steps instances

#stochastic trend - not constant over time() - only adf test can be used 
adf's null hypothesis is data has unit root and alternative is stationary 


```

### Mann Kendall 

Check for deterministic trend.

```{r}

#Since I have seasonal data I cannot use the simple MannKendall()
#another example of functions that need a ts object
SMKtest <- SeasonalMannKendall(ts_inflow_data[,iHP])
print(summary(SMKtest))

#Use yearly date to run Mann Kendall

```

### Spearman Correlation test 

Also check for deterministic trend, for non-seasonal data only.

```{r}
#Deterministic trend with Spearman Correlation Test
print("Results from Spearman Correlation")


#with cor.test you can get test statistics

```

### Augmented Dickey Fuller (ADF)

Used to detect stochastic trend.

```{r}
#Null hypothesis is that data has a unit root
it can still have a determistic trend. you dont need to differenciate 
print("Results for ADF test")

#Now let's try the yearly data
print("Results for ADF test on yearly data/n")

```

##Exercise

Run the stationarity tests for the other variables HP2, HP3, ...

